Analysis and results

### 2023-05-16 analysis/03_analysis.Rmd
Updated 03_analysis to handle MNL as well, including saving draws in-line

# Model calibration

After selecting the model we'd like to run, model calibration proceeds as follows. Note that model calibration is run in a separate R script given how long it takes to run each model.

```{r model-calibration, eval=FALSE}
intercept <- 1                            # Intercept-only
interactions <- 0                         # Include interactions with the intercept-only model
public_affairs <- 0                       # Public affairs
political_ideology <- 0                   # Political ideology
social_views <- 0                         # Social views
charity_voluntarism <- 0                  # Charity and voluntarism
demographics <- 0                         # Demographics
public_political <- 0                     # Public affairs + Political ideology
public_political_social <- 0              # + Social views
public_political_social_charity <- 0      # + Charity and voluntarism
public_political_social_charity_demo <- 0 # + Demographics

# Load data and design
final_data <- read_rds(here::here("data", "derived_data", "final_data.rds"))
X <- read_rds(here::here("data", "derived_data", "final_design.rds"))
if (interactions == 1) X <- read_rds(here::here("data", "derived_data", "final_interactions.rds"))

# Restructure choice data Y
Y <- final_data %>%
  select(contains("Q4")) %>%
  as.matrix()

# Restructure covariate matrix Z
if (intercept == 1) {
  Z <- matrix(data = 1, nrow = dim(X)[1], ncol = 1)
}
if (public_affairs == 1) {
  Z <- final_data %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7 # Public affairs
    )
}
if (political_ideology == 1) {
  Z <- final_data %>%
    select(
      Q5.2 # Political ideology
    )
}
if (social_views == 1) {
  Z <- final_data %>%
    select(
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10 # Social views
    )
}
if (charity_voluntarism == 1) {
  Z <- final_data %>%
    select(
      Q2.7, Q2.8, Q2.5, Q2.6, Q2.9, Q2.10, Q5.4, Q5.5, Q5.3_1:Q5.3_10 # Charity and voluntarism
    )
}
if (demographics == 1) {
  Z <- final_data %>%
    select(
      Q5.12, Q5.13, Q5.14, Q5.15, Q5.16_1:Q5.16_6, Q5.17 # Demographics
    )
}
if (public_political == 1) {
  Z <- final_data %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7, # Public affairs
      Q5.2                                   # Political ideology
    )
}
if (public_political_social == 1) {
  Z <- final_data %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7, # Public affairs
      Q5.2,                                  # Political ideology
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10         # Social views
    )
}
if (public_political_social_charity == 1) {
  Z <- final_data %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7,                           # Public affairs
      Q5.2,                                                            # Political ideology
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10,                                  # Social views
      Q2.7, Q2.8, Q2.5, Q2.6, Q2.9, Q2.10, Q5.4, Q5.5, Q5.3_1:Q5.3_10  # Charity and voluntarism
    )
}
if (public_political_social_charity_demo == 1) {
  Z <- final_data %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7,                           # Public affairs
      Q5.2,                                                            # Political ideology
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10,                                  # Social views
      Q2.7, Q2.8, Q2.5, Q2.6, Q2.9, Q2.10, Q5.4, Q5.5, Q5.3_1:Q5.3_10, # Charity and voluntarism
      Q5.12, Q5.13, Q5.14, Q5.15, Q5.16_1:Q5.16_6, Q5.17               # Demographics
    )
}

@@

# Specify the data for calibration in a list
data <- list(
  R = dim(X)[1],      # Number of respondents
  S = dim(X)[2],      # Number of choice tasks
  A = dim(X)[3],      # Number of choice alternatives
  I = dim(X)[4],      # Number of (estimable) attribute levels
  J = ncol(Z),        # Number of respondent-level covariates

@@

# Run the model and save data and model output
fit <- stan(
  file = here::here("src", "stan_files", "hmnl_noncentered.stan"),
  data = data,
  iter = 10000,
  thin = 5,
  seed = 1234
)
run <- list(data = data, fit = fit)
# if (intercept == 1 & interactions == 0) write_rds(run, here::here("data", "raw_data", "model_runs", "intercept.rds"))
if (intercept == 1 & interactions == 0) write_rds(run, here::here("data", "raw_data", "model_runs", "index_intercept.rds"))
if (intercept == 1 & interactions == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "interactions_5e.rds"))
if (public_affairs == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "public_affairs.rds"))
if (political_ideology == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "political_ideology.rds"))
if (social_views == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "social_views.rds"))
if (charity_voluntarism == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "charity_voluntarism.rds"))
if (demographics == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "demographics.rds"))
if (public_political == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "public_political.rds"))
if (public_political_social == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "public_political_social.rds"))
if (public_political_social_charity == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "public_political_social_charity.rds"))
if (public_political_social_charity_demo == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "public_political_social_charity_demo.rds"))
```

@@

# Compute model fit and extract draws

The following analysis follows specifying the model we'd like to evaluate. Note that the output of the model calibration is too large to push to GitHub, but we can push the draws once we extract them.

## Load model output

```{r load-data-output, warning=FALSE, message=FALSE, eval=FALSE}
# Load all model output
intercept_run <- read_rds(here::here("data", "raw_data", "model_runs", "intercept.rds"))
public_affairs_run <- read_rds(here::here("data", "raw_data", "model_runs", "public_affairs.rds"))
political_ideology_run <- read_rds(here::here("data", "raw_data", "model_runs", "political_ideology.rds"))
social_views_run <- read_rds(here::here("data", "raw_data", "model_runs", "social_views.rds"))
charity_voluntarism_run <- read_rds(here::here("data", "raw_data", "model_runs", "charity_voluntarism.rds"))
demographics_run <- read_rds(here::here("data", "raw_data", "model_runs", "demographics.rds"))
public_political_run <- read_rds(here::here("data", "raw_data", "model_runs", "public_political.rds"))
public_political_social_run <- read_rds(here::here("data", "raw_data", "model_runs", "public_political_social.rds"))
public_political_social_charity_run <- read_rds(here::here("data", "raw_data", "model_runs", "public_political_social_charity.rds"))
public_political_social_charity_demo_run <- read_rds(here::here("data", "raw_data", "model_runs", "public_political_social_charity_demo.rds"))
```

## Compute and save model fit

```{r compute-fit, eval=FALSE}
# Compute fit using PSIS-LOO
intercept_loo <- loo(intercept_run$fit, save_psis = TRUE)
public_affairs_loo <- loo(public_affairs_run$fit, save_psis = TRUE)
political_ideology_loo <- loo(political_ideology_run$fit, save_psis = TRUE)
social_views_loo <- loo(social_views_run$fit, save_psis = TRUE)
charity_voluntarism_loo <- loo(charity_voluntarism_run$fit, save_psis = TRUE)
demographics_loo <- loo(demographics_run$fit, save_psis = TRUE)
public_political_loo <- loo(public_political_run$fit, save_psis = TRUE)
public_political_social_loo <- loo(public_political_social_run$fit, save_psis = TRUE)
public_political_social_charity_loo <- loo(public_political_social_charity_run$fit, save_psis = TRUE)
public_political_social_charity_demo_loo <- loo(public_political_social_charity_demo_run$fit, save_psis = TRUE)

# Compare model fit
loo_compare(intercept_loo, public_affairs_loo, political_ideology_loo, social_views_loo, charity_voluntarism_loo, demographics_loo, public_political_loo, public_political_social_loo, public_political_social_charity_loo, public_political_social_charity_demo_loo)
```

```{r save-fit, eval=FALSE}
# Create data frame
model_fit <- tibble(model = rep(NA, 10), elpd_loo = rep(NA, 10))

# Extract PSIS-LOO fit
model_fit$model[1] <- "intercept"; model_fit$loo[1] <- intercept_loo$elpd_loo
model_fit$model[2] <- "public_affairs"; model_fit$loo[2] <- public_affairs_loo$elpd_loo
model_fit$model[3] <- "political_ideology"; model_fit$loo[3] <- political_ideology_loo$elpd_loo
model_fit$model[4] <- "social_views"; model_fit$loo[4] <- social_views_loo$elpd_loo
model_fit$model[5] <- "charity_voluntarism"; model_fit$loo[5] <- charity_voluntarism_loo$elpd_loo
model_fit$model[6] <- "demographics"; model_fit$loo[6] <- demographics_loo$elpd_loo
model_fit$model[7] <- "public_political"; model_fit$loo[7] <- public_political_loo$elpd_loo
model_fit$model[8] <- "public_political_social"; model_fit$loo[8] <- public_political_social_loo$elpd_loo
model_fit$model[9] <- "public_political_social_charity"; model_fit$loo[9] <- public_political_social_charity_loo$elpd_loo
model_fit$model[10] <- "public_political_social_charity_demo"; model_fit$loo[10] <- public_political_social_charity_demo_loo$elpd_loo

# Save model output
write_rds(model_fit, here::here("analysis", "output", "tables", "model_fit.rds"))
```

## Load model fit objects and extract and save draws

```{r extract-draws, eval=FALSE}
intercept <- 1                            # Intercept-only
interactions <- 1                         # Include interactions with the intercept-only model
public_affairs <- 0                       # Public affairs
political_ideology <- 0                   # Political ideology
social_views <- 0                         # Social views
charity_voluntarism <- 0                  # Charity and voluntarism
demographics <- 0                         # Demographics
public_political <- 0                     # Public affairs + Political ideology
public_political_social <- 0              # + Social views
public_political_social_charity <- 0      # + Charity and voluntarism
public_political_social_charity_demo <- 0 # + Demographics

# Load and extract data and model output
if (intercept == 1 & interactions == 0) run <- read_rds(here::here("data", "raw_data", "model_runs", "intercept.rds"))
if (intercept == 1 & interactions == 1) run <- read_rds(here::here("data", "raw_data", "model_runs", "interactions_5d.rds"))
if (public_affairs == 1) run <- read_rds(here::here("data", "raw_data", "model_runs", "public_affairs.rds"))
if (political_ideology == 1) run <- read_rds(here::here("data", "raw_data", "model_runs", "political_ideology.rds"))
if (social_views == 1) run <- read_rds(here::here("data", "raw_data", "model_runs", "social_views.rds"))
if (charity_voluntarism == 1) run <- read_rds(here::here("data", "raw_data", "model_runs", "charity_voluntarism.rds"))
if (demographics == 1) run <- read_rds(here::here("data", "raw_data", "model_runs", "demographics.rds"))
if (public_political == 1) run <- read_rds(here::here("data", "raw_data", "model_runs", "public_political.rds"))
if (public_political_social == 1) run <- read_rds(here::here("data", "raw_data", "model_runs", "public_political_social.rds"))
if (public_political_social_charity == 1) run <- read_rds(here::here("data", "raw_data", "model_runs", "public_political_social_charity.rds"))
if (public_political_social_charity_demo == 1) run <- read_rds(here::here("data", "raw_data", "model_runs", "public_political_social_charity_demo.rds"))
data <- run$data
fit <- run$fit

# Extract posterior draws
draws <- fit %>% 
  spread_draws(Gamma[i, j]) %>% 
  ungroup() %>% 
  filter(.iteration > 500) %>% 
  rename(
    i = j,
    j = i
  )

# Save draws
if (intercept == 1 & interactions == 0) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "intercept.rds"))
if (intercept == 1 & interactions == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "interactions_5d.rds"))
if (public_affairs == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "public_affairs.rds"))
if (political_ideology == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "political_ideology.rds"))
if (social_views == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "social_views.rds"))
if (charity_voluntarism == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "charity_voluntarism.rds"))
if (demographics == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "demographics.rds"))
if (public_political == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "public_political.rds"))
if (public_political_social == 1)  write_rds(draws, here::here("data", "raw_data", "posterior_draws", "public_political_social.rds"))
if (public_political_social_charity == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "public_political_social_charity.rds"))
if (public_political_social_charity_demo == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "public_political_social_charity_demo.rds"))
```

>>>>

# HMNL model calibration

After selecting the model we'd like to run, model calibration proceeds as follows.

```{r hmnl-model-calibration, eval=FALSE}
intercept <- 1                            # Intercept-only
interactions <- 0                         # Include interactions with the intercept-only model
public_affairs <- 0                       # Public affairs
political_ideology <- 0                   # Political ideology
social_views <- 0                         # Social views
charity_voluntarism <- 0                  # Charity and voluntarism
demographics <- 0                         # Demographics
public_political <- 0                     # Public affairs + Political ideology
public_political_social <- 0              # + Social views
public_political_social_charity <- 0      # + Charity and voluntarism
public_political_social_charity_demo <- 0 # + Demographics

# Load data and design
data <- read_rds(here::here("data", "derived_data", "data.rds"))
X <- read_rds(here::here("data", "derived_data", "hmnl_dummy_design.rds"))
if (interactions == 1) X <- read_rds(here::here("data", "derived_data", "hmnl_dummy_interactions.rds"))

# Restructure choice data Y
Y <- data %>%
  select(contains("Q4")) %>%
  as.matrix()

# Restructure covariate matrix Z
if (intercept == 1) {
  Z <- matrix(data = 1, nrow = dim(X)[1], ncol = 1)
}
if (public_affairs == 1) {
  Z <- data %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7 # Public affairs
    )
}
if (political_ideology == 1) {
  Z <- data %>%
    select(
      Q5.2 # Political ideology
    )
}
if (social_views == 1) {
  Z <- data %>%
    select(
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10 # Social views
    )
}
if (charity_voluntarism == 1) {
  Z <- data %>%
    select(
      Q2.7, Q2.8, Q2.5, Q2.6, Q2.9, Q2.10, Q5.4, Q5.5, Q5.3_1:Q5.3_10 # Charity and voluntarism
    )
}
if (demographics == 1) {
  Z <- data %>%
    select(
      Q5.12, Q5.13, Q5.14, Q5.15, Q5.16_1:Q5.16_6, Q5.17 # Demographics
    )
}
if (public_political == 1) {
  Z <- data %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7, # Public affairs
      Q5.2                                   # Political ideology
    )
}
if (public_political_social == 1) {
  Z <- data %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7, # Public affairs
      Q5.2,                                  # Political ideology
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10         # Social views
    )
}
if (public_political_social_charity == 1) {
  Z <- data %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7,                           # Public affairs
      Q5.2,                                                            # Political ideology
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10,                                  # Social views
      Q2.7, Q2.8, Q2.5, Q2.6, Q2.9, Q2.10, Q5.4, Q5.5, Q5.3_1:Q5.3_10  # Charity and voluntarism
    )
}
if (public_political_social_charity_demo == 1) {
  Z <- data %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7,                           # Public affairs
      Q5.2,                                                            # Political ideology
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10,                                  # Social views
      Q2.7, Q2.8, Q2.5, Q2.6, Q2.9, Q2.10, Q5.4, Q5.5, Q5.3_1:Q5.3_10, # Charity and voluntarism
      Q5.12, Q5.13, Q5.14, Q5.15, Q5.16_1:Q5.16_6, Q5.17               # Demographics
    )
}

@@

# Specify the data for calibration in a list
data_list <- list(
  R = dim(X)[1],      # Number of respondents
  S = dim(X)[2],      # Number of choice tasks
  A = dim(X)[3],      # Number of choice alternatives
  I = dim(X)[4],      # Number of (estimable) attribute levels
  J = ncol(Z),        # Number of respondent-level covariates

@@

# Run the model and save data and model output
fit <- stan(
  file = here::here("src", "stan_files", "hmnl_noncentered.stan"),
  data = data_list,
  iter = 10000,
  thin = 5,
  seed = 1234
)
run <- list(data = data_list, fit = fit)
if (intercept == 1 & interactions == 0) write_rds(run, here::here("data", "raw_data", "model_runs", "intercept.rds"))
if (intercept == 1 & interactions == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "interactions.rds"))
if (public_affairs == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "public_affairs.rds"))
if (political_ideology == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "political_ideology.rds"))
if (social_views == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "social_views.rds"))
if (charity_voluntarism == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "charity_voluntarism.rds"))
if (demographics == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "demographics.rds"))
if (public_political == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "public_political.rds"))
if (public_political_social == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "public_political_social.rds"))
if (public_political_social_charity == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "public_political_social_charity.rds"))
if (public_political_social_charity_demo == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "public_political_social_charity_demo.rds"))

# Extract and save posterior draws
draws <- fit %>% 
  spread_draws(Gamma[i, j]) %>% 
  ungroup() %>% 
  filter(.iteration > 500) %>% 
  rename(
    i = j,
    j = i
  )

# Save draws
if (intercept == 1 & interactions == 0) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "intercept.rds"))
if (intercept == 1 & interactions == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "interactions.rds"))
if (public_affairs == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "public_affairs.rds"))
if (political_ideology == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "political_ideology.rds"))
if (social_views == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "social_views.rds"))
if (charity_voluntarism == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "charity_voluntarism.rds"))
if (demographics == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "demographics.rds"))
if (public_political == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "public_political.rds"))
if (public_political_social == 1)  write_rds(draws, here::here("data", "raw_data", "posterior_draws", "public_political_social.rds"))
if (public_political_social_charity == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "public_political_social_charity.rds"))
if (public_political_social_charity_demo == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "public_political_social_charity_demo.rds"))
```

@@

# MNL model calibration experimental alternative

```{r mnl-model-calibration, eval=FALSE}
################################################################################
# A flat MNL is an experimental alternative for dealing with interactions.
################################################################################

interactions <- 0 # Include interactions

# Load data and design
data <- read_rds(here::here("data", "derived_data", "data.rds"))
X <- read_rds(here::here("data", "derived_data", "mnl_dummy_design.rds"))
if (interactions == 1) X <- read_rds(here::here("data", "derived_data", "mnl_dummy_interactions.rds"))

# Restructure choice data Y
Y <- data %>%
  select(contains("Q4")) %>%
  pivot_longer(contains("Q4"), names_to = "values") %>%
  select(value) %>%
  as.matrix() %>%
  as.vector()

# Specify the data for calibration in a list
data_list <- list(
  N = dim(X)[1],      # Number of observations
  A = dim(X)[2],      # Number of choice alternatives
  L = dim(X)[3],      # Number of (estimable) attribute levels
  Y = Y,              # Vector of observations
  X = X               # Array of observation-level covariates.
)

# Run the model and save data and model output
fit <- stan(
  file = here::here("src", "stan_files", "mnl.stan"),
  data = data_list,
  iter = 10000,
  thin = 5,
  seed = 1234
)
run <- list(data = data_list, fit = fit)
write_rds(run, here::here("data", "raw_data", "model_runs", "mnl.rds"))
if (interactions == 1) write_rds(run, here::here("data", "raw_data", "model_runs", "mnl_interactions.rds"))

# Extract and save posterior draws
draws <- fit %>% 
  spread_draws(Gamma[i, j]) %>% 
  ungroup() %>% 
  filter(.iteration > 500) %>% 
  rename(
    i = j,
    j = i
  )

# Save draws
write_rds(draws, here::here("data", "raw_data", "posterior_draws", "mnl.rds"))
if (interactions == 1) write_rds(draws, here::here("data", "raw_data", "posterior_draws", "mnl_interactions.rds"))
```

The model is calibrated using a multinomial logit model. Each model passed diagnostics checks, including the absence of divergent transitions, as part of model calibration.

```{stan mnl, output.var="mnl", eval=FALSE}
// Index values, observations, and covariates.
data {
  int<lower = 1> N;                  // Number of observations.
  int<lower = 2> A;                  // Number of choice alternatives.
  int<lower = 1> L;                  // Number of (estimable) attribute levels.

  int<lower = 1, upper = A> Y[N];    // Vector of observations.
  matrix[A, L] X[N];                 // Array of observation-level covariates.
}

// Parameters for the multinomial logit.
parameters {
  vector[L] beta;                    // Vector of aggregate beta coefficients.
}

// Multinomial logit model.
model {
  // Standard normal prior for beta.
  beta ~ normal(0, 5);

  // Multinomial logit.
  for (n in 1:N) {
    Y[n] ~ categorical_logit(X[n] * beta);
  }
}

// Quantities conditioned on parameter draws.
generated quantities {
  // vector[N] Y_new;   // Vector of predicted choices.
  vector[N] log_lik; // Log likelihood to estimate loo.

  // Multinomial logit draws and log likelihood per observation.
  for (n in 1:N) {
    log_lik[n] = categorical_logit_lpmf(Y[n] | X[n] * beta);
  }
}
```
